---
title: "word2vec_discourse_viz"
output: html_document
author: "Claire Bergey and Zoe Marshall"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)

library(knitr)
library(here)
library(dplyr)
library(entropy)
library(tidyverse)
library(childesr)
library(RColorBrewer)
library(plotly)
library(feather)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_set(theme_classic(base_size = 14))
```

```{r load-data}
convodata <- read_feather(here("data/childes_convo_reduced_dims.feather")) %>%
  mutate(transcript_id = as.numeric(transcript_id), utterance_order = as.numeric(utterance_order),
         target_child_age = as.numeric(target_child_age)) %>%
  filter(gloss != "yyy", gloss != "xxx")
```

```{r first-transcript}
convodata %>%
  filter(transcript_id == 5465) %>%
  ggplot(aes(V1, V2, color = speaker_code)) +
  geom_point()
```

Above: The first transcript of kid A at 21 months (1.75 years) old.


```{r across-age}
convodata %>%
  filter(target_child_id == 3617) %>%
  ggplot(aes(V1, V2, color = speaker_code)) + 
  facet_grid(~target_child_age) +
  geom_point() 
```

Above: All transcripts from kid A, faceted over age.


```{r last-transcript}
convodata %>%
  filter(transcript_id == 5505) %>%
  ggplot(aes(V1, V2, color = speaker_code)) + 
  geom_point() 
```

Above: The last transcript of kid A at 40 months (3.3 years) old.


```{r last-transcript-3d}
fig <- convodata %>%
  filter(transcript_id == 5505, utterance_order < 300) %>%
  plot_ly(x=~V1, y=~V2, z=~utterance_order, type="scatter3d", 
          mode="markers", 
          text = ~gloss,
          marker = list(size = 3), 
          color=~speaker_code)
fig
```

Above: The last transcript of kid A, with utterance order (time) on the third dimension.


```{r kid-b}
fig <- convodata %>%
  filter(transcript_id == 8625) %>%
  plot_ly(x=~V1, y=~V2, z=~utterance_order, type="scatter3d", 
          mode="markers", 
          text = ~gloss,
          marker = list(size = 3), 
          color=~speaker_code)
fig

```

Above: A transcript from kid B, 3 years old.


```{r kid-a-first}
fig <- convodata %>%
  filter(transcript_id == 5465) %>%
  plot_ly(
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )
fig %>% add_text(textposition = "top right") %>%
  animation_opts(750, redraw = FALSE)
```

Above: The first transcript from kid A, playable over time.


```{r kid-a-last}
fig <- convodata %>%
  filter(transcript_id == 5505) %>%
  plot_ly(
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )
fig %>% add_text(textposition = "top right") %>%
  animation_opts(750, redraw = FALSE)
```

Above: The last transcript of kid A, playable over time.

