---
title: "word2vec_discourse_viz"
output: html_document
author: "Claire Bergey and Zoe Marshall"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)

library(knitr)
library(here)
library(dplyr)
library(entropy)
library(tidyverse)
library(childesr)
library(RColorBrewer)
library(plotly)
library(feather)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_set(theme_classic(base_size = 14))
```

```{r load-data}
convodata <- read_feather(here("data/childes_convo_reduced_dims.feather")) %>%
  mutate(transcript_id = as.numeric(transcript_id), utterance_order = as.numeric(utterance_order),
         target_child_age = as.numeric(target_child_age)) %>%
  filter(gloss != "yyy", gloss != "xxx")
```

```{r}
stop_words <- c('yeah', 'i', 'me', 'my', 'myself', 'we', 'our', 'ours', 'ourselves', 'you', "you're", "you've", "you'll", "you'd", 'your', 'yours', 'yourself', 'yourselves', 'he', 'him', 'his', 'himself', 'she', "she's", 'her', 'hers', 'herself', 'it', "it's", 'its', 'itself', 'they', 'them', 'their', 'theirs', 'themselves', 'what', 'which', 'who', 'whom', 'this', 'that', "that'll", 'these', 'those', 'am', 'is', 'are', 'was', 'were', 'be', 'been', 'being', 'have', 'has', 'had', 'having', 'do', 'does', 'did', 'doing', 'a', 'an', 'the', 'and', 'but', 'if', 'or', 'because', 'as', 'until', 'while', 'of', 'at', 'by', 'for', 'with', 'about', 'against', 'between', 'into', 'through', 'during', 'before', 'after', 'above', 'below', 'to', 'from', 'up', 'down', 'in', 'out', 'on', 'off', 'over', 'under', 'again', 'further', 'then', 'once', 'here', 'there', 'when', 'where', 'why', 'how', 'all', 'any', 'both', 'each', 'few', 'more', 'most', 'other', 'some', 'such', 'no', 'nor', 'not', 'only', 'own', 'same', 'so', 'than', 'too', 'very', 's', 't', 'can', 'will', 'just', 'don', "don't", 'should', "should've", 'now', 'd', 'll', 'm', 'o', 're', 've', 'y', 'ain', 'aren', "aren't", 'couldn', "couldn't", 'didn', "didn't", 'doesn', "doesn't", 'hadn', "hadn't", 'hasn', "hasn't", 'haven', "haven't", 'isn', "isn't", 'ma', 'mightn', "mightn't", 'mustn', "mustn't", 'needn', "needn't", 'shan', "shan't", 'shouldn', "shouldn't", 'wasn', "wasn't", 'weren', "weren't", 'won', "won't", 'wouldn', "wouldn't", 'ha', 'yep')


l <- length(stop_words)

stopless <- convodata

while (l > 0) {
  stopless <- filter(stopless, gloss != stop_words[l])
  l <- l - 1
}
```

```{r first-transcript}
convodata %>%
  filter(transcript_id == 5465) %>%
  ggplot(aes(V1, V2, color = speaker_code)) +
  geom_point()
```

Above: The first transcript of kid A at 21 months (1.75 years) old.


```{r across-age}
convodata %>%
  filter(target_child_id == 3617) %>%
  ggplot(aes(V1, V2, color = speaker_code)) + 
  facet_grid(~target_child_age) +
  geom_point() 
```

Above: All transcripts from kid A, faceted over age.


```{r last-transcript}
convodata %>%
  filter(transcript_id == 5505) %>%
  ggplot(aes(V1, V2, color = speaker_code)) + 
  geom_point() 
```

Above: The last transcript of kid A at 40 months (3.3 years) old.


```{r last-transcript-3d}
fig <- convodata %>%
  filter(transcript_id == 5505, utterance_order < 300) %>%
  plot_ly(x=~V1, y=~V2, z=~utterance_order, type="scatter3d", 
          mode="markers", 
          text = ~gloss,
          marker = list(size = 3), 
          color=~speaker_code)
fig
```

Above: The last transcript of kid A, with utterance order (time) on the third dimension.


```{r kid-b}
fig <- convodata %>%
  filter(transcript_id == 8625) %>%
  plot_ly(x=~V1, y=~V2, z=~utterance_order, type="scatter3d", 
          mode="markers", 
          text = ~gloss,
          marker = list(size = 3), 
          color=~speaker_code)
fig

```

Above: A transcript from kid B, 3 years old.


```{r kid-a-first}
accumulate_by <- function(dat, var) {
  var <- lazyeval::f_eval(var, dat)
  lvls <- plotly:::getLevels(var)
  dats <- lapply(seq_along(lvls), function(x) {
    cbind(dat[var %in% lvls[seq(1, x)], ], frame = lvls[[x]])
  })
  dplyr::bind_rows(dats)
}

fig <- convodata %>%
  filter(transcript_id == 5465) %>%
  
  plot_ly(
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )

fig %>% add_text(textposition = "top right") %>%
  animation_opts(750, redraw = FALSE) %>%
  animation_slider(currentvalue = list(prefix = "Utterance Number ")) 

```

Above: The first transcript from kid A, playable over time.


```{r kid-a-last}
fig <- convodata %>%
  filter(transcript_id == 5505) %>%
  plot_ly(
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )
fig %>% add_text(textposition = "top right") %>%
  animation_opts(750, redraw = FALSE)
```

Above: The last transcript of kid A, playable over time.

```{r kid-a-by-age}
child1 <- convodata %>%
  filter(target_child_id == 3617 & near(target_child_age, 21.72324, tol = 0.5))

child2 <- convodata %>%
  filter(target_child_id == 3617 & near(target_child_age, 26.55907, tol = 0.5))
  
child3 <- convodata %>%
  filter(target_child_id == 3617 & near(target_child_age, 31.16491, tol = 0.5))

child4 <- convodata %>%
  filter(target_child_id == 3617 & near(target_child_age, 35.72353, tol = 0.5))

child5 <- convodata %>%
  filter(target_child_id == 3617 & near(target_child_age, 40.78934, tol = 0.5))

fig1 <-
  plot_ly(
    child1,
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )
fig2 <-
  plot_ly(
    child2,
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )
fig3 <-
  plot_ly(
    child3,
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  ) 
fig4 <-
  plot_ly(
    child4,
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )
fig5 <-
  plot_ly(
    child5,
    x = ~V1,
    y = ~V2,
    color = ~speaker_code,
    frame = ~utterance_order,
    hoverinfo = "text",
    type = 'scatter',
    text = ~gloss,
    mode = 'markers'
  )

fig5
#fig5 %>% add_text(textposition = "top right") %>%
#  animation_opts(750, redraw = FALSE) %>%
#  animation_slider(currentvalue = list(prefix = "Utterance Number ")) 

```