---
title: "Clumpiness"
output: html_document
author: "Claire Bergey"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)

library(knitr)
library(here)
library(dplyr)
library(entropy)
library(tidyverse)
library(childesr)
library(RColorBrewer)
library(plotly)
library(tidyboot)
library(feather)
library(widyr)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_set(theme_classic(base_size = 14))
```

```{r load-data}
convodata <- read_feather(here("data/childes_convo_reduced_dims.feather")) %>%
  mutate(transcript_id = as.numeric(transcript_id), utterance_order = as.numeric(utterance_order),
         target_child_age = as.numeric(target_child_age)) %>%
  filter(target_child_age <= 60) %>%
  filter(!str_detect(gloss, "yyy"),!str_detect(gloss, "xxx"),
         !is.na(target_child_age)) %>%
  mutate(speaker_code = ifelse(speaker_code == "MOM", "MOT", speaker_code),
         speaker_code = ifelse(speaker_code == "DAD", "FAT", speaker_code))
```

```{r lagdata}
lagdata <- convodata %>%
  arrange(transcript_id, utterance_order) %>%
  group_by(transcript_id) %>%
  mutate(dist = sqrt((V1 - lag(V1))^2 + (V2 - lag(V2))^2),
         dist = ifelse(is.na(dist), 0, dist),
         last_speaker = ifelse(lag(speaker_code) == speaker_code, "ME", "YOU"),
         last_speaker = ifelse(is.na(last_speaker), "ME", last_speaker)) %>%
  ungroup()

consecutive_pairwise <- lagdata %>%
  group_by(transcript_id, speaker_code) %>%
  mutate(variance = var(dist),
         variance = ifelse(is.na(variance), 0, variance)) %>%
  ungroup()

consecutive_pairwise %>%
  ggplot(aes(x = target_child_age, y = dist, color = speaker_code)) +
  geom_point(alpha = 0.1, size = 0.2) +
  geom_smooth()
```

Above: Conversational exchange pairwise distances over child age.

```{r dist_exchange}
consecutive_pairwise %>%
  ggplot(aes(x = target_child_age, y = dist, color = speaker_code)) +
  facet_wrap(~last_speaker) +
  geom_point(alpha = 0.1, size = 0.2) +
  geom_smooth()
```

Above: Conversational exchange pairwise distances over child age, faceted by whether I'm replying to myself ("ME") or to someone else ("YOU").

```{r variance}
consecutive_pairwise %>%
  ggplot(aes(x = target_child_age, y = variance, color = speaker_code)) +
  geom_point(alpha = 0.3, size = 0.4) +
  geom_smooth()
```

Above: Conversational exchange pairwise distance variance over child age.

```{r variance_exchange}
consecutive_pairwise %>%
  ggplot(aes(x = target_child_age, y = variance, color = speaker_code)) +
  facet_wrap(~last_speaker) +
  geom_point(alpha = 0.3, size = 0.4) +
  geom_smooth() 
```

Above: Conversational exchange pairwise distance variance over child age, faceted by whether I'm replying to myself ("ME") or to someone else ("YOU").

