---
title: "clustering over distributions of clusters"
author: "Claire Bergey"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(here)
library(dplyr)
library(broom)
library(tidyverse)
library(RColorBrewer)
library(plotly)
library(tidyboot)
library(purrr)
library(tibble)
library(feather)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)

theme_set(theme_classic(base_size = 14))
```

```{r load-data}
convodata <- read_feather(here("data/childes_convo_reduced_dims.feather")) %>%
  mutate(transcript_id = as.numeric(transcript_id),
         utterance_order = as.numeric(utterance_order),
         target_child_age = as.numeric(target_child_age)) %>%
  filter(target_child_age <= 60) %>%
  filter(!str_detect(gloss, "yyy"),!str_detect(gloss, "xxx"),
         !is.na(target_child_age)) %>%
  mutate(speaker_code = ifelse(speaker_code == "MOM", "MOT", speaker_code),
         speaker_code = ifelse(speaker_code == "DAD", "FAT", speaker_code))
```

```{r clustering}
set.seed(111)
all_clusters <- convodata %>%
  select(V1,V2) %>%
  kmeans(12)

all_convo_length <- length(all_clusters$cluster)

convodata$cluster <- all_clusters$cluster
```

```{r wide-recluster}
convodata_normed <- convodata %>%
  count(transcript_id, cluster, name = "num_cluster") %>%
  group_by(transcript_id) %>%
  mutate(num_utts = sum(num_cluster), norm_cluster = num_cluster/num_utts) %>%
  ungroup() 

convodata_normed_clusters <- convodata_normed %>%
  select(-num_cluster, -num_utts) %>%
  pivot_wider(names_from = cluster, values_from = norm_cluster, 
              names_prefix = "cluster_", values_fill = list(norm_cluster = 0))


to_metacluster <- convodata_normed_clusters %>%
  select(-transcript_id)

set.seed(111)
k.max <- 50
wss <- sapply(1:k.max, 
              function(k){kmeans(to_metacluster, k, nstart=50,iter.max = 15 )$tot.withinss})

plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```

Above: total within-cluster sum of squares for clustering over utterance clusters, plotted across # of clusters.

```{r scramble}
convodata_normed_scramble <- convodata %>%
  mutate(scramble_cluster = sample(cluster, replace=FALSE)) %>%
  count(transcript_id, scramble_cluster, name = "num_cluster") %>%
  group_by(transcript_id) %>%
  mutate(num_utts = sum(num_cluster), norm_cluster = num_cluster/num_utts) %>%
  ungroup() %>%
  select(-num_cluster, -num_utts) %>%
  pivot_wider(names_from = scramble_cluster, values_from = norm_cluster, 
              names_prefix = "cluster_", values_fill = list(norm_cluster = 0))

to_scramble_cluster <- convodata_normed_scramble %>%
  select(-transcript_id)

scramble_wss <- sapply(1:k.max, 
              function(k){kmeans(to_scramble_cluster, k, 
                                 nstart=50,iter.max = 15 )$tot.withinss})

plot(1:k.max, scramble_wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```

Above: total within-cluster sum of squares clustering over scrambled utterances.

```{r}
scrambled_metaclustering <- to_scramble_cluster %>%
  kmeans(15)

convodata_normed_clusters$scrambled_metaclusters <- scrambled_metaclustering$cluster
```

```{r}
clusters_over_clusters <- to_metacluster %>%
  kmeans(15)

convodata_normed_clusters$metaclusters <- clusters_over_clusters$cluster

convodata_metaclusters <- convodata %>%
  left_join(convodata_normed_clusters %>% select(transcript_id, 
                                                 metaclusters, scrambled_metaclusters), 
            by = "transcript_id") %>%
  left_join(convodata_normed %>% select(transcript_id, cluster, norm_cluster), 
            by = c("transcript_id","cluster"))
  
```

```{r}
convodata_metaclusters %>%
  group_by(transcript_id) %>% 
  summarise(metaclusters = first(metaclusters), 
            age = first(target_child_age),
            id = first(target_child_id)) %>%
  ggplot(aes(x = metaclusters)) +
  geom_histogram()
```

Above: distribution of metaclusters among real transcripts.

```{r}
convodata_metaclusters %>%
  group_by(transcript_id) %>% 
  summarise(scrambled_metaclusters = first(scrambled_metaclusters), 
            age = first(target_child_age),
            id = first(target_child_id)) %>%
  ggplot(aes(x = scrambled_metaclusters)) +
  geom_histogram()
```

Above: distribution of metaclusters among scrambled utterance transcripts.

```{r}
convodata_metaclusters %>%
  ggplot(aes(x = cluster)) +
  facet_wrap(~metaclusters) +
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]), bins = 12) +
  labs(y = "Density normalized by metacluster")
```

Above: distribution of utterance clusters in each metacluster.

```{r}
convodata_metaclusters %>%
  group_by(transcript_id) %>% 
  summarise(metaclusters = first(metaclusters), 
            age = first(target_child_age),
            id = first(target_child_id)) %>%
  ggplot(aes(x = age)) +
  facet_wrap(~metaclusters) +
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]), bins = 12) +
  labs(y = "Density normalized by metacluster")

```

Above: distribution of ages in each metacluster.

```{r}


```