---
title: "markov-models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(here)
library(feather)
library(tidyverse)
library(plotly)
library(entropy)
library(tidyboot)
library(purrr)
library(DT)
library(markovchain)
library(reshape)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)
```

```{r read-data}
childes_htmm <- read_feather(here("data/childes_htmm_25.feather"))
childes_dict <- read_csv(here("data/childes_gloss_word_dict.csv"))
childes_htmm_hmm <- read_csv(here("data/childes_htmm_hmm.csv"))
```

```{r markovmodel}
seq <- childes_htmm_hmm %>%
  mutate(max_topic = if_else(nchar(max_topic) == 1, paste0("0", max_topic),
                            paste0(max_topic))) 

htmm_mm <- markovchainFit(seq$max_topic)

htmm_tmatrix <- as.data.frame(htmm_mm$estimate@transitionMatrix)

hmm_mm <- markovchainFit(seq$reconstruction)

hmm_tmatrix <- as.data.frame(hmm_mm$estimate@transitionMatrix)
```

```{r heatmap-htmm}

htmm_matrix_long <- htmm_tmatrix %>% rownames_to_column() %>%
  melt() %>%
  dplyr::rename(current_state = rowname, next_state = variable)

htmm_matrix_long %>%
  ggplot(aes(x=current_state, y=next_state, fill=value)) + 
  geom_tile()
```

Transition matrix heatmap of HTMM max likelihood utterance topics among 3 y/os.

```{r heatmap-hmm}

hmm_matrix_long <- hmm_tmatrix %>% rownames_to_column() %>%
  melt() %>%
  dplyr::rename(current_state = rowname, next_state = variable)

hmm_matrix_long %>%
  ggplot(aes(x=current_state, y=next_state, fill=value)) + 
  geom_tile()

```

Transition matrix heatmap of HMM utterance hidden states among 3 y/os.


```{r}

childes_htmm_hmm <- childes_htmm_hmm %>%
  mutate(lead_max_topic = lead(max_topic),
         lead_speaker_code = lead(speaker_code))

squished_seq <- childes_htmm_hmm %>%
  dplyr::filter(!(lead_max_topic == max_topic & 
                    lead_speaker_code == speaker_code)) %>%
  mutate(max_topic = if_else(nchar(max_topic) == 1, paste0("0", max_topic),
                            paste0(max_topic))) 

squished_htmm_mm <- markovchainFit(squished_seq$max_topic)

squished_htmm_tmatrix <- as.data.frame(squished_htmm_mm$estimate@transitionMatrix)

squished_htmm_matrix_long <- squished_htmm_tmatrix %>% rownames_to_column() %>%
  melt() %>%
  dplyr::rename(current_state = rowname, next_state = variable)

squished_htmm_matrix_long %>%
  ggplot(aes(x=current_state, y=next_state, fill=value)) + 
  geom_tile()
```

Transition matrix heatmap of HTMM max likelihood utterance topics among 3 y/os after repeated topics produced by the same person are squished.


```{r}
childes_htmm <- childes_htmm %>%
  mutate(lead_max_topic = lead(max_topic),
         lead_speaker_code = lead(speaker_code))

seq_all <- childes_htmm %>%
  mutate(max_topic = if_else(nchar(max_topic) == 1, paste0("0", max_topic),
                            paste0(max_topic)))

squished_seq_all <- childes_htmm %>%
  dplyr::filter(!(lead_max_topic == max_topic & 
                    lead_speaker_code == speaker_code)) %>%
  mutate(max_topic = if_else(nchar(max_topic) == 1, paste0("0", max_topic),
                            paste0(max_topic))) 

all_mm <- markovchainFit(seq_all$max_topic)
all_tmatrix <- as.data.frame(all_mm$estimate@transitionMatrix)
squished_all_mm <- markovchainFit(squished_seq_all$max_topic)
squished_all_tmatrix <- as.data.frame(squished_all_mm$estimate@transitionMatrix)

all_tmatrix %>% rownames_to_column() %>%
  reshape::melt() %>%
  dplyr::rename(current_state = rowname, next_state = variable) %>%
  ggplot(aes(x=current_state, y=next_state, fill=value)) + 
  geom_tile()

```

Transition matrix heatmap of HTMM max likelihood utterance topics in all of CHILDES.

```{r squished-all}

squished_all_tmatrix %>% rownames_to_column() %>%
  reshape::melt() %>%
  dplyr::rename(current_state = rowname, next_state = variable) %>%
  ggplot(aes(x=current_state, y=next_state, fill=value)) + 
  geom_tile()

```

Transition matrix heatmap of HTMM max likelihood utterance topics in all of CHILDES after repeated topics produced by the same person are squished.

```{r squished_all_hist}
seq_all %>%
  ggplot() +
  geom_histogram(aes(x = lead_max_topic, binwidth = 1)) +
  facet_wrap(~max_topic)

squished_seq_all %>%
  ggplot() +
  geom_histogram(aes(x = lead_max_topic, binwidth = 1)) +
  facet_wrap(~max_topic)
```

```{r zeroed}

zero <- squished_seq_all %>%
  filter(lead_max_topic != max_topic) %>%
  mutate(lead_max_topic = lead(max_topic)) %>%
  filter(lead_max_topic != max_topic)
zero_mm <- markovchainFit(zero$max_topic)
zero_tmatrix <- as.data.frame(zero_mm$estimate@transitionMatrix)


zero_tmatrix %>% rownames_to_column() %>%
  reshape::melt() %>%
  dplyr::rename(current_state = rowname, next_state = variable) %>%
  ggplot(aes(x=current_state, y=next_state, fill=value)) + 
  geom_tile()

zero_hmm <- childes_htmm_hmm %>%
  mutate(lead_state = lead(reconstruction)) %>%
  filter(lead_state != reconstruction)

zero_hmm_mm <- markovchainFit(zero_hmm$reconstruction)

```