---
title: "topics analyses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(here)
library(feather)
library(tidyverse)
library(tidyboot)
library(purrr)
library(gsl)
library(DT)
library(reshape)
library(markovchain)
library(gsl)
library(reticulate)
use_condaenv("r-reticulate")

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = TRUE, tidy = FALSE)
```

```{r read-data}
childes_htmm_pdwz <- read_feather(here("data/childes_htmm_all_models"))
dict <- read_csv(here("data/childes_gloss_word_dict.csv"))
```

```{r entropy-par-child}

childes_htmm_pdwz <- childes_htmm_pdwz %>%
  mutate(age_in_months = floor(target_child_age)) %>%
  mutate(speaker_code = if_else(str_detect("CHI", speaker_code), 
                                "CHI", speaker_code))

topic_count_vector <- childes_htmm_pdwz %>%
  group_by(max_topic_15, transcript_id, age_in_months, speaker_code) %>%
  count() %>% ungroup()

counts <- topic_count_vector %>%
  mutate(name = paste0(transcript_id, speaker_code)) 

counts <- split(counts$n, counts$name)

num_topics <- n_distinct(childes_htmm_pdwz$max_topic_15)

```

```{python}
import ndd
k_val = r.num_topics
entropies_dict = {}

for key, value in r.counts.items():
  entropies_dict[key] = ndd.entropy(value, k = k_val)
```

```{r}
topic_entropies <- py$entropies_dict %>%
  as.data.frame() %>% t() %>% as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(name = rowname, entropy = V1) %>%
  mutate(name = str_remove(name, "X"))

topic_entropy <- topic_count_vector %>% 
  mutate(name = paste0(transcript_id, speaker_code)) %>%
  left_join(topic_entropies, by = "name")

par_child_entropies <- topic_entropy %>%
  group_by(transcript_id, speaker_code) %>%
  summarise(entropy = first(entropy), age = first(age_in_months)) %>%
  group_by(age, speaker_code) %>%
  mutate(entropy_random = entropy[sample(row_number())]) %>%
  ungroup()

```

```{r graph_entropies_over_age}
topic_entropy %>%
  filter(speaker_code == "CHI" | speaker_code == "MOT") %>%
  ggplot(aes(age_in_months, entropy, color = speaker_code)) +
  geom_point(alpha = 0.08) +
  geom_smooth() +
  ylab("Entropy") +
  xlab("Age") +
  ggtitle("Entropy of Topics") +
  theme_classic()
```

Mother and child topic entropies over age.

```{r graph_random_entropies}
par_child_entropies %>%
  pivot_longer(cols = c(entropy, entropy_random), names_to = "is_random", values_to = "entropy") %>%
  pivot_wider(values_from = entropy, names_from = speaker_code) %>%
  ggplot(aes(x = MOT, y = CHI, color = age)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~is_random)
```

Correlation between mother and child entropies (left) and the same correlation when speakers' entropies are swapped within speaker identity (among mothers, among children) within a 1-month age range (right).

```{r types-of-utterances}
childes_htmm_pdwz %>%
  filter(type != "broken for coding", type != "question exclamation", 
         type != "quotation precedes") %>%
  ggplot(aes(x = max_topic_15)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~type, scale = "free")

```

CHILDES has some codes for type of utterance that are really coarse-grained--declarative, interruption, imperative emphatic, etc. Do our topics predict these these utterance types? 

```{r branching-out}
childes_htmm_pdwz %>%
  dplyr::filter(speaker_code == "CHI" | speaker_code == "MOT") %>%
  ggplot(aes(x = age_in_months, fill = as.factor(max_topic_15))) +
  geom_bar(position = "fill") +
  facet_wrap(~speaker_code) +
  theme_classic() 
```

How do parents and children change their topic use over time? How do children branch out into using new kinds of utterances?

